\textbf{Transparency to hybrid open access through publisher-provided metadata: An article-level study of Elsevier}

### 2021-02-04 analysis/manuscript.Rmd
minor tex fixes for binder

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: no
    theme: readable
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = ""
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
    } else{
      return(x)
    }
  }
)
```

@@

Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%. As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas. In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}). This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles. This order mostly reflects the disciplines' overall publication output. According to the Open Science Monitor by the European Commission, the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.

>>>>

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
csl: apa.csl
bibliography: references.bib
---

```{r knitr_options, , echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300,
  fig.pos = "H", 
  out.extra = "",
  dev= c("png", "cairo_ps")
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "f", digits = 1, drop0trailing = TRUE))
    } else{
      return(x)
    }
  }
)
```

@@

Over the last five years, Elsevier recorded growth in the uptake of hybrid OA for each of our three indicators: The number of hybrid OA articles published per year doubled, the number of hybrid journals with at least one OA article grew by 21%, and the share of hybrid OA articles relative to closed-accessed articles in these journals increased from 2.6% to 3.7%. As @Laakso_2016, we found only a weak relationship between citation metrics like the SNIP and hybrid OA uptake and observed differences in the uptake of hybrid OA among journal subject areas. In particular, we found the highest count of hybrid OA articles in physical sciences journals (see Table \ref{tab:subject_area_table}). This was followed by the life sciences and health sciences, whereas the social sciences had the lowest count of hybrid articles. This order mostly reflects the disciplines' overall publication output. According to the Open Science Monitor by the European Commission [@OS_Monitor], the physical sciences and mathematics publish the most articles, followed by the health sciences, life sciences, and social sciences and humanities.

### 2021-02-04 analysis/manuscript.tex
minor tex fixes for binder

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\title{Transparency to hybrid open access invoicing through publisher-provided metadata: A longitudinal article-level study of Elsevier}
\shorttitle{Hybrid OA Elsevier}
\author{Najko Jahn}
\affiliation{State and University Library Göttingen, University of Göttingen, Germany}
\author{Lisa Matthias}
\affiliation{}
\author{Mikael Laakso}

\abstract{The financial flows in scholarly journal publishing have become increasingly complex with the growth of open access (OA), and the lack of data and transparency into these flows has been persistent. Opaqueness has concerned especially hybrid OA, where subscription-based journals also publish OA content if an optional fee is paid. By leveraging Elsevier’s recently introduced article metadata concerning invoiced recipients, this study provides the first longitudinal publisher-level study of uptake and invoicing of hybrid OA articles. The results show that the share of OA articles in Elseviers hybrid journals has grown slowly, from 2.6\% of all articles on average in 2015 to 3.7\% in 2019. About a third of hybrid OA articles are invoiced directly through an agreement with research funders or other type of organisational sponsor (e.g. national library consortia), with the majority invoiced to authors. 83\% of agreement-invoiced hybrid OA articles are to European organisations, suggesting a strong international skewness. The study discovered a large discrepancy in choice of publication license between author-invoiced articles (dominantly CC BY-NC-ND) and agreement-invoiced articles (dominantly CC BY). The results demonstrate the value of publisher-provided metadata covering invoiced recipients of articles, which would be even more useful through wider adoption among publishers.
}

\begin{document}

\maketitle

@@

Over the last five years, Elsevier recorded growth in the uptake of
hybrid OA for each of our three indicators: The number of hybrid OA
articles published per year doubled, the number of hybrid journals with
at least one OA article grew by 21\%, and the share of hybrid OA
articles relative to closed-accessed articles in these journals
increased from 2.6\% to 3.7\%. As Laakso \& Björk (2016), we found only
a weak relationship between citation metrics like the SNIP and hybrid OA
uptake and observed differences in the uptake of hybrid OA among journal
subject areas. In particular, we found the highest count of hybrid OA
articles in physical sciences journals (see Table
\ref{tab:subject_area_table}). This was followed by the life sciences
and health sciences, whereas the social sciences had the lowest count of
hybrid articles. This order mostly reflects the disciplines' overall
publication output. According to the Open Science Monitor by the
European Commission, the physical sciences and mathematics publish the
most articles, followed by the health sciences, life sciences, and
social sciences and humanities.

@@

@@

\leavevmode\hypertarget{ref-McNutt_2019}{}%
McNutt, M. (2019). Opinion: \\
textquotedblleftPlan s\\
textquotedblright falls short for society publishers\\
textemdashand for the researchers they serve. \emph{Proceedings of the
National Academy of Sciences}, \emph{116}(7), 2400--2403.
\url{https://doi.org/10.1073/pnas.1900359116}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
% make captions italic

@@

\title{\textbf{Transparency to hybrid open access  through publisher-provided metadata: An article-level study of Elsevier}}
\shorttitle{Hybrid OA Elsevier}
\threeauthors{Najko Jahn}{Lisa Matthias}{Mikael Laakso}
\threeaffiliations{Göttingen State and University Library, University of Göttingen\\
Platz der Göttinger Sieben 1, 37073 Göttingen, Germany\\
najko.jahn@sub.uni-goettingen.de
}{Department of Political Science, John F. Kennedy Institute, Freie Universität Berlin \\
Lansstraße 7-9, 14195 Berlin, Germany \\
l.a.matthia@gmail.com
}{Information Systems Science, Hanken School of Economics \\
Arkadiankatu 22, 00100 Helsinki, Finland \\
mikael.laakso@hanken.fi
}

\authornote{Correspondence concerning this article should be addressed to Najko Jahn}
% \affiliation{State and University Library Göttingen, University of Göttingen, Germany}
% \author{Lisa Matthias}
% \affiliation{tba}
% \author{Mikael Laakso}
% \affiliation{tba}
%

\abstract{With the growth of open access (OA), the financial flows in scholarly journal
publishing have become increasingly complex, but comprehensive data and
transparency into these flows are still lacking. The opaqueness is especially
concerning for hybrid OA, where subscription-based journals publish individual
articles as OA if an optional fee is paid. This study addresses the lack of
transparency by leveraging Elsevier article metadata and provides the first
longitudinal publisher-level study of hybrid OA uptake and invoicing. Our
results show that Elsevier's hybrid OA uptake has grown steadily but slowly from
2015-2019, doubling the number of hybrid OA articles published per year and
increasing the share of OA articles in Elsevier's hybrid journals from 2.6\% to
3.7\% of all articles. Further, we find that most hybrid OA articles were
invoiced directly to authors, followed by articles invoiced through agreements
with research funders, institutions, or consortia, with only a few funding
bodies driving hybrid OA uptake. As such, our findings point to the role of
publishing agreements and OA policies in hybrid OA publishing. Our results
further demonstrate the value of publisher-provided metadata improve the
transparency in scholarly publishing by linking invoicing data to bibliometrics.}

\begin{document}
\maketitle

@@

Over the last five years, Elsevier recorded growth in the uptake of
hybrid OA for each of our three indicators: The number of hybrid OA
articles published per year doubled, the number of hybrid journals with
at least one OA article grew by 21\%, and the share of hybrid OA
articles relative to closed-accessed articles in these journals
increased from 2.6\% to 3.7\%. As Laakso \& Björk (2016), we found only
a weak relationship between citation metrics like the SNIP and hybrid OA
uptake and observed differences in the uptake of hybrid OA among journal
subject areas. In particular, we found the highest count of hybrid OA
articles in physical sciences journals (see Table
\ref{tab:subject_area_table}). This was followed by the life sciences
and health sciences, whereas the social sciences had the lowest count of
hybrid articles. This order mostly reflects the disciplines' overall
publication output. According to the Open Science Monitor by the
European Commission (European Commission, 2019), the physical sciences
and mathematics publish the most articles, followed by the health
sciences, life sciences, and social sciences and humanities.

@@

\leavevmode\hypertarget{ref-OS_Monitor}{}%
European Commission. (2019). \emph{Open science monitor: Trends for open
access to publications}.
\url{https://ec.europa.eu/info/research-and-innovation/strategy/goals-research-and-innovation-policy/open-science/open-science-monitor/trends-open-access-publications_en}

@@

\leavevmode\hypertarget{ref-McNutt_2019}{}%
McNutt, M. (2019). Opinion: Plan S falls short for society
publishers--and for the researchers they serve. \emph{Proceedings of the
National Academy of Sciences}, \emph{116}(7), 2400--2403.
\url{https://doi.org/10.1073/pnas.1900359116}
